---
title: "mortality_modeling"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      prompt = FALSE,
                      error = TRUE,
                      message = FALSE,
                      cache = FALSE,
                      eval = TRUE,
                      eval.after = "fig.cap")

library(tidyverse)
library(rFIA)
library(ggplot2)
library(ggExtra)
library(rgdal)
library(sp)
library(ggsci)
library(raster)
library(splines)
library(lme4)
library(patchwork)
select <- dplyr::select

theme_set(theme_bw())
theme_update(text = element_text(size=16, color = "black"),
             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(),
             strip.background = element_blank(),
             panel.border=element_rect(size=1.5))

# source("/Users/DanielPerret/Box/01. daniel.perret Workspace/status_trends/new_rFIA_fns.R")
# source("/Users/DanielPerret/Box/01. daniel.perret Workspace/status_trends/old_rFIA_fns.R")
source("/Users/DanielPerret/Box/01. daniel.perret Workspace/status_trends/growMort_rewrite.R")

```

This is just an exploratory document that builds and plays with some of the climateXdisturbance mortality models that I've been building and thinking about recently

The structure that I've kind of landed on after the last week or so of thinking is such
- response -- proportional mortality over remeasurement period at each site
- fixed effects:
- climate anomalies (MAT, MAP, CMD, etc)
- proportion of *ecoregion* impacted by diseases, insects, fire
- interactions between all of above
- random effects:
- site in ecoregion nested intercepts
- ecoregion slopes?

So this should be a GLMM with a binomial distribution and weights proportional to the denominator of the response.
(is that right???)



BUILDING COMBINED DATAFRAME:

Here's the area estimation for disturbance agents:

```{r}
## playing with area estimation based on either condition classes or whether mortality was observed

popinfo <- rFIA::getDesignInfo(all.fia,type="ALL") %>% 
  left_join(all.fia$POP_STRATUM %>% 
              select(STRATUM_CN=CN,
                     EXPNS),
            by="STRATUM_CN")

View(plot.summary)

fire.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19,
         agent_key=="fire") %>% 
  pull(PLT_CN) %>% 
  unique()

insect.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19,
         agent_key=="insect") %>% 
  pull(PLT_CN) %>% 
  unique()

disease.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19,
         agent_key=="disease") %>% 
  pull(PLT_CN) %>% 
  unique()

animal.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19,
         agent_key=="animal") %>% 
  pull(PLT_CN) %>% 
  unique()

competition.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19,
         agent_key=="competition") %>% 
  pull(PLT_CN) %>% 
  unique()

weather.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19,
         agent_key=="weather") %>% 
  pull(PLT_CN) %>% 
  unique()

land.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19,
         agent_key=="land_use") %>% 
  pull(PLT_CN) %>% 
  unique()

unknown.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19,
         agent_key=="unknown") %>% 
  pull(PLT_CN) %>% 
  unique()

all.plots <- all.fia$TREE %>% 
  filter(most.recent=="yes",
         SPCD==19) %>% 
  pull(PLT_CN) %>% 
  unique()

mort.area.all <- all.fia$PLOT %>% 
  left_join(popinfo %>% 
              select(PLT_CN,EXPNS)) %>%
  summarise(fire.area = sum(EXPNS*ifelse(PLT_CN %in% fire.plots, 1, 0), na.rm=T),
            insect.area = sum(EXPNS*ifelse(PLT_CN %in% insect.plots, 1, 0), na.rm=T),
            disease.area = sum(EXPNS*ifelse(PLT_CN %in% disease.plots, 1, 0), na.rm=T),
            id.area = sum(EXPNS*ifelse(PLT_CN %in% c(disease.plots,insect.plots), 1, 0), na.rm=T),
            fid.area = sum(EXPNS*ifelse(PLT_CN %in% c(fire.plots,disease.plots,insect.plots), 1, 0), na.rm=T),
            animal.area = sum(EXPNS*ifelse(PLT_CN %in% animal.plots, 1, 0), na.rm=T),
            competition.area = sum(EXPNS*ifelse(PLT_CN %in% competition.plots, 1, 0), na.rm=T),
            weather.area = sum(EXPNS*ifelse(PLT_CN %in% weather.plots, 1, 0), na.rm=T),
            land.area = sum(EXPNS*ifelse(PLT_CN %in% land.plots, 1, 0), na.rm=T),
            unknown.area = sum(EXPNS*ifelse(PLT_CN %in% unknown.plots, 1, 0), na.rm=T),
            total.area = sum(EXPNS*ifelse(PLT_CN %in% all.plots, 1, 0), na.rm=T))

mort.area.er <- all.fia$PLOT %>% 
  left_join(popinfo %>% 
              select(PLT_CN,EXPNS)) %>%
  group_by(ECOSUBCD) %>%
  filter(ECOSUBCD %in% er4.abla$MAP_UNIT_S) %>% 
  summarise(area.fire = sum(EXPNS*ifelse(PLT_CN %in% fire.plots, 1, 0), na.rm=T),
            area.insect = sum(EXPNS*ifelse(PLT_CN %in% insect.plots, 1, 0), na.rm=T),
            area.disease = sum(EXPNS*ifelse(PLT_CN %in% disease.plots, 1, 0), na.rm=T),
            area.id = sum(EXPNS*ifelse(PLT_CN %in% c(disease.plots,insect.plots), 1, 0), na.rm=T),
            area.fid = sum(EXPNS*ifelse(PLT_CN %in% c(fire.plots,disease.plots,insect.plots), 1, 0), na.rm=T),
            area.animal = sum(EXPNS*ifelse(PLT_CN %in% animal.plots, 1, 0), na.rm=T),
            area.competition = sum(EXPNS*ifelse(PLT_CN %in% competition.plots, 1, 0), na.rm=T),
            area.weather = sum(EXPNS*ifelse(PLT_CN %in% weather.plots, 1, 0), na.rm=T),
            area.land = sum(EXPNS*ifelse(PLT_CN %in% land.plots, 1, 0), na.rm=T),
            area.unknown = sum(EXPNS*ifelse(PLT_CN %in% unknown.plots, 1, 0), na.rm=T),
            area.total = sum(EXPNS*ifelse(PLT_CN %in% all.plots, 1, 0), na.rm=T))

#get MORT DECOMP
```

Here's getting stand-level characteristics
```{r}

abla.ba.change <- growMort_dlp(db = all.fia,
                               stateVar = "BAA",
                               #polys=er4.abla,
                               treeDomain = SPCD==19,
                               totals = TRUE,
                               byPlot = TRUE,
                               returnSpatial = F, 
                               nCores = 4,
                               variance = TRUE,
                               sizeThresh=1,
                               evals = evals)

all.ba.change <-   growMort_dlp(db = all.fia,
                                stateVar = "BAA",
                                #polys=er4.abla,
                                #treeDomain = SPCD==19,
                                totals = TRUE,
                                byPlot = TRUE,
                                returnSpatial = F, 
                                nCores = 4,
                                variance = TRUE,
                                sizeThresh=1,
                                evals = evals)

ba <- abla.ba.change %>% 
  select(PLT_CN,
         PREV_BAA_abla = PREV_BAA,
         CURR_BAA_abla = CURR_BAA) %>% 
  left_join(all.ba.change %>% 
              select(PLT_CN,
                     PREV_BAA_all = PREV_BAA,
                     CURR_BAA_all = CURR_BAA),
            by="PLT_CN") %>% 
  mutate(TOT_BAA_CHNG_all = CURR_BAA_all - PREV_BAA_all,
         TOT_BAA_CHNG_abla = CURR_BAA_abla - PREV_BAA_abla,
         BAA_CHNG_PROP_all = TOT_BAA_CHNG_all/PREV_BAA_all,
         BAA_CHNG_PROP_abla = TOT_BAA_CHNG_abla/PREV_BAA_abla)

```


And here's the complete dataset -- all mortality estimates and climate data at the plot level, disturbance information at the ecoregion subsection level

```{r}

## this is from mort_decomp

evals <- c(21903,41903,61903,81903,161903,301903,321903,351903,411903,491903,531903,561903)

tpa.change <- all.fia %>% 
  growMort_dlp(db = .,
               stateVar = "TPA",
               #polys=er4.abla,
               treeDomain = SPCD==19,
               totals = TRUE,
               byPlot = TRUE,
               returnSpatial = T, 
               nCores = 4,
               variance = TRUE,
               sizeThresh=1,
               evals = evals)

tpa.mort.agent <- all.fia %>% 
  growMort_dlp(db = .,
               stateVar = "TPA",
               #polys=er4.abla,
               treeDomain = SPCD==19,
               totals = TRUE,
               grpBy=agent_key,
               byPlot = TRUE,
               returnSpatial = T, 
               nCores = 4,
               variance = TRUE,
               sizeThresh=1,
               evals = evals)

mort.decomp <- tpa.change %>% 
  sf::st_drop_geometry() %>% 
  filter(CURR_TPA > 0 | PREV_TPA >0) %>% 
  left_join(tpa.mort.agent %>%
              sf::st_drop_geometry() %>% 
              #filter(agent_key=="fire") %>% 
              #filter(agent_key%in%c("insects","disease")) %>% 
              group_by(PLT_CN) %>% 
              summarise(TOT_MORT = sum(CURR_TPA-PREV_TPA,na.rm=T)*-1,
                        FIRE_MORT = sum((CURR_TPA-PREV_TPA)*ifelse(agent_key=="fire",1,0),na.rm=T)*-1,
                        DISEASE_MORT = sum((CURR_TPA-PREV_TPA)*ifelse(agent_key=="disease",1,0),na.rm=T)*-1,
                        INSECT_MORT = sum((CURR_TPA-PREV_TPA)*ifelse(agent_key=="insect",1,0),na.rm=T)*-1,
                        ID_MORT = sum((CURR_TPA-PREV_TPA)*
                                        ifelse(agent_key%in%c("insect","disease"),1,0))*-1),na.rm=T)%>%
  mutate(TOT_MORT = ifelse(is.na(TOT_MORT),0,TOT_MORT),
         TOT_MORTPERC = round(TOT_MORT/PREV_TPA,6),
         FIRE_MORT = ifelse(is.na(FIRE_MORT),0,FIRE_MORT),
         INSECT_MORT = ifelse(is.na(INSECT_MORT),0,INSECT_MORT),
         DISEASE_MORT = ifelse(is.na(DISEASE_MORT),0,DISEASE_MORT),
         ID_MORT = ifelse(is.na(ID_MORT),0,ID_MORT),
         CHNG_TPA_TOT = (CURR_TPA-PREV_TPA),
         RECR_est = (CURR_TPA-PREV_TPA) + TOT_MORT,
         fire.pace = ifelse(FIRE_MORT>RECR_est, "no", "yes"),
         id.pace = ifelse(ID_MORT > RECR_est, "no", "yes"),
         other.pace = ifelse(TOT_MORT-(FIRE_MORT+ID_MORT) > RECR_est, "no", "yes"),
         all.pace = ifelse(TOT_MORT > RECR_est, "no","yes")) %>% 
  left_join(all.fia$PLOT %>% 
              select(PLT_CN, INVYR, ELEV, LON, LAT, ECOSUBCD,   MAT:DD1040_base),
            by="PLT_CN") %>% 
  mutate(MAT_anom = MAT_remper - MAT_base,
         MAP_anom = MAP_remper - MAP_base,
         CMD_anom = CMD_remper - CMD_base,
         MWMT_anom = MWMT_remper - MWMT_base,
         MCMT_anom = MCMT_remper - MCMT_base,
         MSP_anom = MSP_remper - MSP_base,
         AHM_anom = AHM_remper - AHM_base,
         SHM_anom = SHM_remper - SHM_base,
         NFFD_anom = NFFD_remper - NFFD_base,
         FFP_anom = FFP_remper - FFP_base,
         EMT_anom = EMT_remper - EMT_base,
         EXT_anom = EXT_remper - EXT_base,
         CMI_anom = CMI_remper - CMI_base,
         Eref_anom = Eref_remper - Eref_base,
  ) %>% 
  left_join(mort.area.er %>% 
              mutate(area.fire.prop = area.fire/area.total,
                     area.insect.prop = area.insect/area.total,
                     area.disease.prop = area.disease/area.total,
                     area.id.prop = area.id/area.total,
                     area.fid.prop = area.fid/area.total
              ), 
            by="ECOSUBCD") %>% 
  left_join(all.fia$COND %>% 
              filter(CONDID==1) %>% 
              mutate(
                FORTYPGRP_key = case_when(
                  substr(FORTYPCD,1,2) == 20 ~ "Douglas fir",
                  substr(FORTYPCD,1,2) == 22 ~ "Ponderosa pine",
                  substr(FORTYPCD,1,2) == 24 ~ "Western white pine",
                  substr(FORTYPCD,1,2) == 26 ~ "Fir/spruce/hemlock",
                  substr(FORTYPCD,1,2) == 27 ~ "Fir/spruce/hemlock", # check this one!
                  substr(FORTYPCD,1,2) == 28 ~ "Lodgepole pine",
                  substr(FORTYPCD,1,2) == 30 ~ "Hemlock/Sitka spruce",
                  substr(FORTYPCD,1,2) == 32 ~ "Western larch",
                  substr(FORTYPCD,1,2) == 36 ~ "Other western softwoods",
                  substr(FORTYPCD,1,2) == 70 ~ "Elm/ash/cottonwood",
                  substr(FORTYPCD,1,2) == 90 ~ "Aspen/birch",
                  substr(FORTYPCD,1,2) == 91 ~ "Alder/maple",
                  substr(FORTYPCD,1,2) == 97 ~ "Woodland hardwoods",
                  substr(FORTYPCD,1,2) == 99 ~ "Nonstocked",
                  is.na(FORTYPCD) ~ "Unknown"),
                FORTYPCD_key = case_when(
                  FORTYPCD==201 ~ "Douglas fir",
                  FORTYPCD==221 ~ "Ponderosa pine",
                  FORTYPCD==241 ~ "Western white pine",
                  FORTYPCD==261 ~ "White fir",
                  FORTYPCD==262 ~ "Red fir",
                  FORTYPCD==263 ~ "Noble fir",
                  FORTYPCD==264 ~ "Pacific silver fir",
                  FORTYPCD==265 ~ "Engelmann spruce",
                  FORTYPCD==266 ~ "Engelmann spruce/subalpine fir",
                  FORTYPCD==267 ~ "Grand fir",
                  FORTYPCD==268 ~ "Subalpine fir",
                  FORTYPCD==269 ~ "Blue spruce",
                  FORTYPCD==270 ~ "Mountain hemlock",
                  FORTYPCD==271 ~ "Alaska-yellow-cedar",
                  FORTYPCD==281 ~ "Lodgepole pine",
                  FORTYPCD==301 ~ "Western hemlock",
                  FORTYPCD==304 ~ "Western redcedar",
                  FORTYPCD==321 ~ "Western larch",
                  FORTYPCD==366 ~ "Limber pine",
                  FORTYPCD==367 ~ "Whitebark pine",
                  FORTYPCD==368 ~ "Misc. softwoods",
                  FORTYPCD==703 ~ "Cottonwood",
                  FORTYPCD==709 ~ "Cottonwood/willow",
                  FORTYPCD==901 ~ "Aspen",
                  FORTYPCD==911 ~ "Red alder",
                  FORTYPCD==971 ~ "Deciduous oak woodland",
                  FORTYPCD==974 ~ "Cercocarpus woodland",
                  FORTYPCD==975 ~ "Intermountain maple woodland",
                  FORTYPCD==999 ~ "Nonstocked",
                  is.na(FORTYPCD) ~ "Unknown")) %>% 
              select(PLT_CN,STDAGE,STDSZCD,contains("DSTRBCD"),contains("_key")),
            by="PLT_CN") %>% 
  left_join(ba,
            by="PLT_CN") %>% 
  filter(!PLT_CN %in% hold,
         !is.na(TOT_MORTPERC),
         !is.na(STDAGE),
         TOT_MORTPERC>0,
         !is.na(area.id.prop),
         !is.na(MAT_remper)) %>% 
  mutate(ECOSECCD = substr(ECOSUBCD, 1, nchar(ECOSUBCD)-1),
         ECOPRVCD = substr(ECOSUBCD,1,nchar(ECOSUBCD)-2),
         PREV_BA_meanabla = PREV_BAA_abla/PREV_TPA,
         across(MAT:Eref_anom,scale,.names = "{.col}_scaled"),
         STDAGE_scaled=scale(STDAGE),
         across(contains("BAA"),scale,.names="{.col}_scaled"))

```

Here's an initial model:

```{r}
library(ggeffects)
library(performance)

m.er <- glmer(data = mort.decomp %>% 
                filter(!is.na(TOT_MORTPERC),
                       TOT_MORTPERC>0,
                       !is.na(area.id.prop),
                       !is.na(MAT_remper)) %>%
                mutate(across(MAT:Eref_anom,scale,.names = "{.col}_scaled")),
              
              formula = TOT_MORTPERC ~ INVYR +
                PREV_BAA_abla_scaled + PREV_BAA_all_scaled +
                area.fire.prop*MAT_anom_scaled+#*CMD_anom_scaled +
                area.id.prop*MAT_anom_scaled+#*CMD_anom_scaled +
                area.fire.prop:area.id.prop+
                + FORTYPGRP_key*MAT_anom_scaled +
                (1|ECOSUBCD),
              
              # formula = TOT_MORTPERC ~ #INVYR +
              #   area.fire.prop*MAT_anom_scaled +
              #   #PREV_BAA_abla_scaled*PREV_BAA_all_scaled +
              #   area.id.prop*MAT_anom_scaled +
              #   area.id.prop:area.fire.prop +
              #   MAT_anom_scaled*FORTYPGRP_key +
              #   (1|ECOSUBCD),
              
              # formula = TOT_MORTPERC ~ area.fire.prop*MAT_anom +
              #   area.fire.prop*MAP_anom + area.fire.prop*CMD_anom +
              #   area.id.prop*MAT_anom + area.id.prop*MAP_anom +
              #   area.id.prop*CMD_anom +check
              #   (1|ECOSUBCD),
              family = binomial(),
              weights = PREV_TPA,
              nAGQ = 0,
              control = glmerControl(optimizer = "nlminbwrap"))

summary(m.er)

performance(m.er)

plot(m.er@frame$TOT_MORTPERC, 
     predict(m.er,type="response",re.form=NULL));abline(0,1,col="red")

plot(predict(m.er,re.form=NULL,type="response"),predict(m.er,re.form=NA,type="response"))

check_model(m.er)

```

```{r}
ggpredict(m.er,
          terms = c("area.fire.prop [all]", "MAT_anom_scaled [-2,-1,0,1,2]"),
          type="fe") %>% 
  ggplot(.,
         aes(x, predicted, groups = group)) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill="black",
              alpha=0.1)+
  geom_line(aes(col = group),
            lwd=1.5) +
  labs(x = "Area proportion impacted by fire",
       y= "Predicted mortality (marginal)") +
  scale_color_manual(name = "MAT anomaly \n(scaled)",
                     values = c("-2" = "dodgerblue4",
                                "-1" = "dodgerblue1",
                                "0" = "black",
                                "1" = "orange",
                                "2" = "red"))

ggpredict(m.er,
          terms = c("area.id.prop [all]", "MAT_anom_scaled [-2,-1,0,1,2]"),
          type="fe") %>% 
  ggplot(.,
         aes(x, predicted, groups = group)) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill="black",
              alpha=0.1)+
  geom_line(aes(col = group),
            lwd=1.5) +
  labs(x = "Area proportion impacted by insects/disease",
       y= "Predicted mortality (marginal)") +
  scale_color_manual(name = "MAT anomaly \n(scaled)",
                     values = c("-2" = "dodgerblue4",
                                "-1" = "dodgerblue1",
                                "0" = "black",
                                "1" = "orange",
                                "2" = "red"))

```

```{r}
ggpredict(m.er,
          terms = c("MAT_anom_scaled [all]", "area.fire.prop [0, 0.25, 0.5, 0.75, 1]"),
          type="fe") %>% 
  ggplot(.,
         aes(x, predicted, groups = group)) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill="darkgray",
              alpha=0.4)+
  geom_line(aes(col = group),
            lwd=1.5) +
  labs(x = "Scaled MAT anomaly",
       y= "Predicted mortality (marginal)") +
  scale_color_manual(name = "Proportion burned",
                     values = c("0" = "dodgerblue4",
                                "0.25" = "dodgerblue1",
                                "0.5" = "black",
                                "0.75" = "orange",
                                "1" = "red"))

ggpredict(m.er,
          terms = c("MAT_anom_scaled [all]", "area.id.prop [0, 0.25, 0.5, 0.75, 1]"),
          type="fe") %>% 
  ggplot(.,
         aes(x, predicted, groups = group)) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill="darkgray",
              alpha=0.4)+
  geom_line(aes(col = group),
            lwd=1.5) +
  labs(x = "Scaled MAT anomaly",
       y= "Predicted mortality (marginal)") +
  scale_color_manual(name = "Proportion insects/disease",
                     values = c("0" = "dodgerblue4",
                                "0.25" = "dodgerblue1",
                                "0.5" = "black",
                                "0.75" = "orange",
                                "1" = "red"))

```

```{r}

s.mean <- mean(mort.decomp$MAT_anom)
s.sd <- sd(mort.decomp$MAT_anom)

ggpredict(m.er,
          terms = c("MAT_anom.er_scaled [-4:5,by=0.1]", 
                    "area.id.prop [0, 0.3, 0.6, 1]", 
                    "area.fire.prop [0,0.3,0.6,1]"),
          #condition = c(INVYR=2019),
          #               PREV_BAA_abla_scaled = 0,
          #               PREV_BAA_all_scaled = 0),
          type="fe") %>% as.data.frame() %>% 
  ggplot(.,
         aes((x*s.sd)+s.mean, predicted*100, groups = group)) +
  geom_ribbon(aes(ymin = conf.low*100,
                  ymax = conf.high*100),
              fill="darkgray",
              alpha=0.4)+
  geom_line(aes(col = group),
            lwd=1.5) +
  geom_vline(xintercept = s.mean,
             lty=2)+
  labs(x = "MAT anomaly (C)",
       y= "Predicted % mortality") +
  scale_color_manual(name = "Proportion \ninsects/disease",
                     values = c("0" = "dodgerblue4",
                                "0.3" = "dodgerblue1",
                                #"0.5" = "black",
                                "0.6" = "orange",
                                "1" = "red")) +
  geom_rug(sides = "b", data = mort.decomp, aes(x = MAT_anom), inherit.aes=F)+
  facet_wrap(facets=~facet,
             labeller = as_labeller(c("0" = "0% burned",
                                      "0.3" = "30% burned",
                                      "0.6" = "60% burned",
                                      "1" = "100% burned")))

```

```{r}

ggpredict(m.er,
          terms = c("area.id.prop [0:1, by = 0.05]",
                    "MAT_anom_scaled [-2,-1,0,1,2]", 
                    "area.fire.prop [0,0.3,0.6,1]"),
          condition = c(FORTYPGRP_key = "Fir/spruce/hemlock"),
          type="fe") %>%
  ggplot(.,
         aes(x, predicted*100, groups = group)) +
  geom_ribbon(aes(ymin = conf.low*100,
                  ymax = conf.high*100),
              fill="darkgray",
              alpha=0.4)+
  geom_line(aes(col = group),
            lwd=1.5)+
  labs(x = "Area insect/disease",
       y= "Predicted % mortality") +
  scale_color_manual(name = "MAT anomaly \n(scaled)",
                     values = c("-2" = "dodgerblue4",
                                "-1" = "dodgerblue1",
                                "0" = "black",
                                "1" = "orange",
                                "2" = "red")) +
  facet_wrap(facets=~facet,
             labeller = as_labeller(c("0" = "0% burned",
                                      "0.3" = "30% burned",
                                      "0.6" = "60% burned",
                                      "1" = "100% burned")))

```

```{r}
ggpredict(m.er,
          terms = c("STDAGE_scaled [all]","PREV_BAA_all_scaled [-2,-1,0,1,2]"),
          #condition = c(FORTYPGRP_key = "Fir/spruce/hemlock"),
          type="fe") %>% 
  ggplot(.,
         aes(x, 
             predicted,
             groups=group)) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill="darkgray",
              alpha=0.4)+
  geom_line(aes(col=group),
            lwd=1.5) +
  labs(x = "Stand age (scaled)",
       y= "Predicted mortality (marginal)") 

```


```{r}
ggpredict(m.er,
          terms = c("PREV_BAA_all_scaled [all]", "STDAGE_scaled [-2,-1,0,1,2]"),
          #condition = c(FORTYPGRP_key = "Fir/spruce/hemlock"),
          type="fe") %>% 
  ggplot(.,
         aes(x, predicted,
             groups=group)) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill="darkgray",
              alpha=0.4)+
  geom_line(lwd=1.5,
            aes(col=group)) +
  labs(x = "Stand basal area (scaled)",
       y= "Predicted mortality (marginal)") 

```





















